Query Optimization for Dynamic Imputation

### 2017-05-02 rebuttal.tex

\documentclass{vldb}
\usepackage{xcolor}
\newcommand{\resp}[1]{{\color{blue}{#1}}}
\newcommand{\todo}[1]{{\color{red}{\textit{#1}}}}
\newcommand{\ProjectName}[0]{ImputeDB}

### 2017-05-05 rebuttal.tex

\item - S4.4.1: Imputation over the raw data is being used as ground truth. Is there any reason to suspect that this is the correct thing to do? Perhaps joining multiple tables together exposes new attributes that makes imputation *more* reliable? At the very least, it would be useful to understand why there is so much error in $\alpha=0$ on Q2 and Q8.

    \resp{For our purposes, imputation over the raw data is a reasonable ground truth. This
        is because the analyst, in practice, should be considering the tradeoffs on quality
        and performance of using ImputeDB versus a production imputation system on the raw
        data. Unfortunately, they would never have access to a real ground truth in such a
        scenario, and moreover, finding ground truth for use even in an empirical evaluation
        of imputation methods can be challenging because artificially dirtying the data
        often does not capture the complex reasons for missingness in real world datasets.
        We update the discussion in S4.4.1 to emphasize this point.}

    \item - There are a number of opportunities for trimming in the figures, tables, and algorithms: Algorithms 1-3 could be compacted a bunch, Table 1 is full of whitespace (e.g., make it a 3-column {figure*}), Tables 1-2 could also be summarized as CDFs, Figure 6 is full of whitespace, Table 3 could be put queries on a single row (e.g., Q1 would easily fit on 1 line), and the references have a lot of redundant information (URLs, for example).

      \resp{We have compacted the algorithms and tables. Note that Tables 1-2 show the percentage of missing values for \emph{each attribute}, not as an overall percentage, so cannot be summarized as CDFs. We think that the URLs add value to the references, particularly for data sets.}

@@

\item D7. When defining the cost of impute operator, how do you know a priori the number of iterations for the algorithm (given it has to converge)?

\item D8. Are there impute implementation that are non-blocking? Or do they always have to perform multiple passes over the data?

\item D9. Additional experiments I suggest to include:
-- Consider a scenario where the user submits multiple similar queries. Show after how many such queries your approach has no longer a competitive advantage compared to performing imputation on the data.
-- Try different values for $\alpha$ (not just 0 and 1).
-- Show runtime of enumeration algorithm as queries get more complex.

\resp{We explored the planning times for queries with increasing number of joins. The standard algorithm presented in the original version submitted handles
up to 5 joins, with columns requiring imputation in all tables, with planning time of approximately one second. We added a small optimization to help reduce
planning times for a larger number of joins. We maintain an approximation of the Pareto sets, rather than the exact set, by only adding non-dominated plans that
are sufficiently different. This helps reduce some of the increases in planning time for a larger number of joins.}

### 2017-05-05 paper/algorithm.tex
paper: remove reference to missing algorithm
added discussion of planning times to paper and rebuttal

Finally, we note that for the queries we have examined, this exponential blowup does not affect the practical performance of our optimizer.
Recall that the planner maintains different plans for different dirty sets, keeping only those plans that are not dominated by others.
So in many cases we can drop some of the intermediate plans generated at each operator.
The worst case complexity only occurs if the dirty sets tracked are distinct through the entire planning phase.
Planning times are discussed in \Cref{sec:results}.

>>>>

Finally, we note that for the queries we have examined, this exponential blowup does not affect the practical performance of our optimizer.
Recall that the planner maintains different plans for different dirty sets, keeping only those plans that are not dominated by others.
So in many cases we can drop some of the intermediate plans generated at each operator.
The worst case complexity only occurs if the dirty sets tracked are distinct through the entire planning phase.

In order to support a large number of joins, with reasonable planning times, \ProjectName{} can maintain approximations of the Pareto sets,
rather than exact sets. To do so, dominated plans are still removed based on new non-dominated plans, but new plans are only added
if they are significantly different from any existing plan in the set. The distance between plans is determined based on the cost penalty and time
values associated with a plan.  Planning times are discussed further in \Cref{sec:results}.

### 2017-05-05 paper/experiments.tex
added discussion of planning times to paper and rebuttal

>>>>

\subsubsection{Planning times}
\begin{table}
\centering
\input{tables/planning_summary_exact_pareto.tex}
\caption{Planning times are sub-second for up to 4 joins. For 5+  joins planning times rise. Many practical queries
can be formulated with a smaller number of joins, but to support more complex queries \ProjectName{} provides an
optimization that maintains approximate Pareto sets for queries with many joins. This reduces planning time
for queries with 7 joins by approximately 45\%, on average.}
\label{table:planning-times}
\end{table}

We explored the practical implications of our planning algorithm by constructing a series of queries with increasing number of joins.
We used the CDC tables and evaluated a combination of self-joins and joins across reference tables. Planning times for queries
involving 1 to 5 joins, which although limited is a practical value for real world exploratory queries, take less than a second.
For queries with 7 joins, planning time increases substantially to 29 seconds, on average. We introduced an optimization that reduces
the plan space by maintaining only approximations of the Pareto set for queries involving a large number of joins.
With this optimization, the same set of queries have an average planning time of 16 seconds (2 seconds SE).
